---
title: "Monte Carlo Simulation Study for Spatial Spillover Synthetic Control"
author: "Replication Package"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

# Overview

This vignette reproduces the finite‐sample evaluation of the spatial spillover synthetic control estimator.  
We conduct a Monte Carlo study under controlled data‐generating processes (DGPs) with spatial dependence and interference.

The following functions are assumed to be already installed and available (part of the replication package):

- `run_many_sim()`: Run repeated simulations under one scenario  
- `summarize_many()`: Summarize bias, RMSE, and 95% coverage  
- `make_rook_W()`: Generate rook‐adjacency matrix on an m×n grid  
- `run_scenario_many()`: Convenience wrapper for one scenario  
- `mc_grid_study()`: Grid simulation across (N, T0, ρ) values  

```{r setup, message=FALSE, warning=FALSE}
rm(list = ls())
library(scspill)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(knitr)
library(kableExtra)

source("R/01_utils.R")
source("R/10_sc_spillover.R")
source("R/21_mcmc_alpha.R")
source("R/22_mcmc_sar.R")
source("R/30_simulation_.R")
source("R/40_geweke_latest.R")
source("R/41_robustness_check.R")
Rcpp::sourceCpp("src/20_mcmc.cpp")
Rcpp::sourceCpp("src/40_geweke_latest.cpp")

# Assume the simulation engine functions are already part of the package
set.seed(20251030)
```

# Data Generating Process

We simulate an untreated pre‐period of length $begin:math:text$ T_0 $end:math:text$ and a post‐period of length $begin:math:text$ T_1 $end:math:text$.  
Units lie on a spatial lattice, and outcomes follow a spatial autoregressive structure with spillover from treated units.

Key features of the DGP:

- Spatial correlation governed by $begin:math:text$ \\rho $end:math:text$
- Spillover weights $begin:math:text$ w $end:math:text$ and adjacency matrix $begin:math:text$ W $end:math:text$
- One or more covariates (dimension $begin:math:text$K$end:math:text$)
- A fixed set of treated units
- MCMC estimation with diffuse priors

# Running the Monte Carlo Study

The command below runs multiple scenarios over a grid of $begin:math:text$ (N, T_0, \\rho) $end:math:text$.  
Each scenario involves repeated sampling and estimation.

```{r}
make_rook_W <- function(nrow, ncol, normalize = FALSE) {
  stopifnot(nrow >= 1L, ncol >= 1L)
  idx <- matrix(seq_len(nrow * ncol), nrow, ncol, byrow = TRUE)
  N <- nrow * ncol
  W <- matrix(0, N, N)

  for (r in 1:nrow) {
    for (c in 1:ncol) {
      i <- idx[r, c]
      if (r > 1) {
        W[i, idx[r - 1, c]] <- 1
      }
      if (r < nrow) {
        W[i, idx[r + 1, c]] <- 1
      }
      if (c > 1) {
        W[i, idx[r, c - 1]] <- 1
      }
      if (c < ncol) W[i, idx[r, c + 1]] <- 1
    }
  }

  if (normalize) {
    rs <- rowSums(W)
    W <- W / pmax(rs, 1) # 0割回避
  }
  W
}

run_scenario_many <- function(
  n_sims,
  dgp_args,
  seeds = NULL,
  M = 5000,
  burn = 2000,
  step_rho = 0.02
) {
  res <- run_many_sim(
    n_sims = n_sims,
    dgp_args = dgp_args,
    seeds = seeds,
    M = M,
    burn = burn,
    step_rho = step_rho
  )
  list(
    summary = summarize_many(res),
    raw = res
  )
}

mc_grid_study <- function(
  Ns,
  T0s,
  rhos,
  T1,
  sims_per = 100,
  K = 1,
  beta = c(1.0),
  sigma2 = 1.0,
  treated_idx = 1:4,
  M = 5000,
  burn = 2000,
  step_rho = 0.02,
  seeds = NULL
) {
  # グリッド展開
  grid <- expand.grid(N = Ns, T0 = T0s, rho = rhos, stringsAsFactors = FALSE)

  # 乱数種
  if (is.null(seeds)) {
    seeds <- sample.int(.Machine$integer.max, nrow(grid) * sims_per)
  }
  # シナリオごとの種をスライス
  get_seeds_i <- function(i) {
    idx <- ((i - 1L) * sims_per + 1L):(i * sims_per)
    seeds[idx]
  }

  # 実行
  details <- vector("list", nrow(grid))
  summaries <- vector("list", nrow(grid))

  for (i in seq_len(nrow(grid))) {
    N <- grid$N[i]
    T0 <- grid$T0[i]
    rho <- grid$rho[i]

    # N は正方格子前提（論文の体裁を踏襲）。
    m <- round(sqrt(N))
    if (m * m != N) {
      stop(sprintf(
        "N=%d は完全平方数ではありません（%dx%dにできません）。",
        N,
        m,
        m
      ))
    }

    # W と w
    W <- make_rook_W(m, m)
    w <- numeric(N)
    w[intersect(treated_idx, seq_len(N))] <- 1

    # 真の α
    alpha_true <- rep(0, N)
    alpha_true[1] <- 0.5
    alpha_true[2] <- -0.2
    alpha_true[3:4] <- 0.4
    alpha_true[5:10] <- 0.1 / 6

    # DGP 引数
    dgp_args <- list(
      T0 = T0,
      T1 = T1,
      N = N,
      W = W,
      w = w,
      rho = rho,
      sigma2 = sigma2,
      alpha = alpha_true,
      K = K,
      beta = beta
    )

    # 多回実行
    process_time <- system.time({
      out <- run_scenario_many(
        n_sims = sims_per,
        dgp_args = dgp_args,
        seeds = get_seeds_i(i),
        M = M,
        burn = burn,
        step_rho = step_rho
      )
    })
    print(process_time)

    # シナリオ情報を列として付与
    smry <- out$summary
    smry$N <- N
    smry$T0 <- T0
    smry$T1 <- T1
    smry$rho <- rho
    # 列順を調整（読みやすさ）
    smry <- smry[, c(
      "N",
      "T0",
      "T1",
      "rho",
      "method",
      "bias_point",
      "rmse_point",
      "cover95_point"
    )]

    details[[i]] <- out$raw
    summaries[[i]] <- smry
  }

  list(
    summary = do.call(rbind, summaries),
    details = details,
    grid = grid
  )
}
```


```{r run-grid}
output_dir <- "inst/tables/mc_result"
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

for (N in c(16, 36, 64)) {
  for (T0 in c(20, 50)) {
    for (T1 in c(20, 50)) {
      study <- mc_grid_study(
        Ns = c(N), # grid size: 4×4 and 6×6
        T0s = c(T0), # pre-period length
        rhos = c(-0.8, -0.3, -0.1, 0.0, 0.1, 0.3, 0.8), # spatial coefficient c(-0.8, -0.3, -0.1, 0.0, 0.1, 0.3, 0.8)
        T1 = T1, # post-treatment length (20 or 50)
        sims_per = 1000, # repetitions per scenario. It takes too long time if set 1000.
        K = 1, # number of covariates
        beta = c(1.0), # true regression slope
        sigma2 = 0.1, # innovation variance
        treated_idx = 1:4, # treated units
        M = 2500, # MCMC iterations
        burn = 500, # burn-in
        step_rho = 0.05 # MH step size
      )
      output_path <- paste0(
        output_dir,
        "/mc_study_N=",
        N,
        "T0=",
        T0,
        "T1=",
        T1,
        ".csv"
      )
      write.csv(study$summary, output_path, row.names = FALSE)
      rm(study)
    }
  }
}

```

The output table contains, for each scenario and estimation method:

- `bias`: Bias of the point estimate
- `rmse`: Root mean squared error
- `cover95`: Coverage of the nominal 95% interval





```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(readr)
library(glue)

## ------------------------------------------------------------
## 1. mc_study_*.csv を読み込んで一つの long データにする
## ------------------------------------------------------------

read_mc_results <- function(dir = "inst/tables/mc_result") {
  files <- list.files(
    dir,
    pattern = "^mc_study_N=\\d+T0=\\d+T1=\\d+\\.csv$",
    full.names = TRUE
  )

  meta <- tibble(file = files) %>%
    mutate(
      fname = basename(file),
      N = as.integer(str_match(fname, "N=(\\d+)")[, 2]),
      T0 = as.integer(str_match(fname, "T0=(\\d+)")[, 2]),
      T1 = as.integer(str_match(fname, "T1=(\\d+)")[, 2]),
      data = map(
        file,
        ~ read_csv(.x, show_col_types = FALSE) %>%
          # csv 側に N/T0/T1 が入っていてもここで落としておく
          select(-any_of(c("N", "T0", "T1")))
      )
    )

  res <- meta %>%
    unnest(cols = data)

  # メトリクスを分かりやすい名前にリネーム
  res <- res %>%
    rename(
      Bias = bias_point,
      RMSE = rmse_point,
      CoverageRate = cover95_point
    )

  res
}

mc_res <- read_mc_results()

## ------------------------------------------------------------
## 2. Table 1 (Bias / RMSE: Proposed, SCM, BSCM) の panel を作る
##    T0, T1 ごとに 1 個の panel を出力
## ------------------------------------------------------------

make_table1_panel <- function(
  res,
  T0_val,
  T1_val,
  file = glue("inst/tables/table1_T0={T0_val}_T1={T1_val}.tex")
) {
  df <- res %>%
    filter(T0 == T0_val, T1 == T1_val) %>%
    filter(method %in% c("Proposed", "SCM", "BSCM")) %>%
    mutate(
      method = factor(method, levels = c("Proposed", "SCM", "BSCM")),
      rho = as.numeric(rho)
    )

  N_vals <- sort(unique(df$N))
  rho_vals <- sort(unique(df$rho))

  # 出力開始
  con <- file(file, open = "w", encoding = "UTF-8")
  on.exit(close(con), add = TRUE)

  # ヘッダ（caption や table 環境は好みに応じて調整）
  cat("\\begin{table}[!htbp]\n", file = con)
  cat("\\centering\n", file = con)
  cat(glue("% Panel: T0 = {T0_val}, T1 = {T1_val}\n"), file = con)
  cat("\\begin{tabular}{l", file = con)
  for (i in seq_along(N_vals)) {
    cat(" rrr", file = con)
  }
  cat("}\n\\toprule\n", file = con)

  # 1 行目: N 見出し
  cat(" & ", file = con)
  cat(
    paste0(
      glue("\\multicolumn{{3}}{{c}}{{$N = {N_vals}$}}"),
      collapse = " & "
    ),
    file = con
  )
  cat(" \\\\\n", file = con)

  # 2 行目: メソッド名
  cat(" & ", file = con)
  cat(
    paste(
      rep(c("Proposed", "SCM", "BSCM"), length(N_vals)),
      collapse = " & "
    ),
    file = con
  )
  cat(" \\\\\n\\midrule\n", file = con)

  fmt <- function(x) formatC(x, format = "f", digits = 3)

  ## ---- Bias ブロック ----
  cat(
    "\\multicolumn{",
    1 + 3 * length(N_vals),
    "}{l}{Bias} \\\\\n",
    sep = "",
    file = con
  )

  for (r in rho_vals) {
    cat(glue("${r}$"), file = con)
    for (N0 in N_vals) {
      tmp <- df %>%
        filter(N == N0, rho == r) %>%
        arrange(method)

      vals <- fmt(tmp$Bias)
      cat(" & ", paste(vals, collapse = " & "), file = con)
    }
    cat(" \\\\\n", file = con)
  }

  cat("\\midrule\n", file = con)

  ## ---- RMSE ブロック ----
  cat(
    "\\multicolumn{",
    1 + 3 * length(N_vals),
    "}{l}{RMSE} \\\\\n",
    sep = "",
    file = con
  )

  for (r in rho_vals) {
    cat(glue("${r}$"), file = con)
    for (N0 in N_vals) {
      tmp <- df %>%
        filter(N == N0, rho == r) %>%
        arrange(method)

      vals <- fmt(tmp$RMSE)
      cat(" & ", paste(vals, collapse = " & "), file = con)
    }
    cat(" \\\\\n", file = con)
  }

  cat("\\bottomrule\n\\end{tabular}\n\\end{table}\n", file = con)
}

## 例：図の Panel (a), (b) に対応しそうなもの
## （必要なら T1_val は適宜変えてください）
make_table1_panel(
  mc_res,
  T0_val = 20,
  T1_val = 20,
  file = "inst/tables/table1_panel_a.tex"
)
make_table1_panel(
  mc_res,
  T0_val = 50,
  T1_val = 20,
  file = "inst/tables/table1_panel_b.tex"
)

## ------------------------------------------------------------
## 3. Table 2 (Coverage Rate: Proposed のみ) を作る
## ------------------------------------------------------------

make_table2_tex <- function(
  mc_res,
  method_proposed = "SCSPILL",
  out_file = "inst/tables/mc_table2.tex"
) {
  # ρ は numeric に揃える
  mc_res <- mc_res %>%
    mutate(rho = as.numeric(rho))

  # proposed method だけ取り出し，カバレッジの MC 平均を計算
  cov_df <- mc_res %>%
    filter(method == method_proposed) %>%
    group_by(T0, N, rho) %>%
    summarise(coverage = mean(CoverageRate, na.rm = TRUE), .groups = "drop")

  # 想定：T0 は {20, 50}，N は {16, 36, 64}，rho は共通
  t0_vals <- sort(unique(cov_df$T0))
  rho_vals <- sort(unique(cov_df$rho))

  # T0 = 20 のブロック
  tab20 <- cov_df %>%
    filter(T0 == t0_vals[1]) %>%
    select(rho, N, coverage) %>%
    pivot_wider(names_from = N, values_from = coverage) %>%
    arrange(rho)

  # T0 = 50 のブロック
  tab50 <- cov_df %>%
    filter(T0 == t0_vals[2]) %>%
    select(rho, N, coverage) %>%
    pivot_wider(names_from = N, values_from = coverage) %>%
    arrange(rho)

  # 列名は N=16,36,64 を想定
  n_cols <- sort(setdiff(names(tab20), "rho"))

  #--------------------------------------------------------
  # 3. LaTeX 文字列を構築
  #--------------------------------------------------------
  lines <- c(
    "\\begin{table}[!htbp]",
    "\\centering",
    "\\caption{Table 2: Coverage Rate of 95\\% Credible Interval for the proposed method}",
    sprintf("%% using method == \"%s\"", method_proposed),
    "\\begin{tabular}{lccc ccc}",
    "\\toprule",
    " & \\multicolumn{3}{c}{$T_0 = 20$} & \\multicolumn{3}{c}{$T_0 = 50$}\\\\",
    "\\cmidrule(lr){2-4} \\cmidrule(lr){5-7}",
    "$\\rho$ & $N = 16$ & $N = 36$ & $N = 64$ & $N = 16$ & $N = 36$ & $N = 64$ \\\\",
    "\\midrule"
  )

  for (r in rho_vals) {
    row20 <- tab20[tab20$rho == r, n_cols, drop = FALSE]
    row50 <- tab50[tab50$rho == r, n_cols, drop = FALSE]

    vals20 <- sprintf("%.3f", as.numeric(row20[1, ]))
    vals50 <- sprintf("%.3f", as.numeric(row50[1, ]))

    line_r <- sprintf(
      "% .1f & %s & %s & %s & %s & %s & %s \\\\",
      r,
      vals20[1],
      vals20[2],
      vals20[3],
      vals50[1],
      vals50[2],
      vals50[3]
    )
    lines <- c(lines, line_r)
  }

  lines <- c(lines, "\\bottomrule", "\\end{tabular}", "\\end{table}")

  writeLines(lines, con = out_file)
  invisible(out_file)
}

mc_res <- read_mc_results("inst/tables/mc_result")
make_table2_tex(mc_res)

```
