---
title: "Monte Carlo Simulation Study for Spatial Spillover Synthetic Control"
author: "Replication Package"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

# Overview

This vignette reproduces the finite‐sample evaluation of the spatial spillover synthetic control estimator.  
We conduct a Monte Carlo study under controlled data‐generating processes (DGPs) with spatial dependence and interference.

The following functions are assumed to be already installed and available (part of the replication package):

- `run_many_sim()`: Run repeated simulations under one scenario  
- `summarize_many()`: Summarize bias, RMSE, and 95% coverage  
- `make_rook_W()`: Generate rook‐adjacency matrix on an m×n grid  
- `run_scenario_many()`: Convenience wrapper for one scenario  
- `mc_grid_study()`: Grid simulation across (N, T0, ρ) values  

```{r setup, message=FALSE, warning=FALSE}
rm(list = ls())
library(scspill)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(knitr)
library(kableExtra)

source("R/01_utils.R")
source("R/10_sc_spillover.R")
source("R/21_mcmc_alpha.R")
source("R/22_mcmc_sar.R")
source("R/30_simulation_.R")
source("R/40_geweke_latest.R")
source("R/41_robustness_check.R")
Rcpp::sourceCpp("src/20_mcmc.cpp")
Rcpp::sourceCpp("src/40_geweke_latest.cpp")

# Assume the simulation engine functions are already part of the package
set.seed(20251030)
```

# Data Generating Process

We simulate an untreated pre‐period of length $begin:math:text$ T_0 $end:math:text$ and a post‐period of length $begin:math:text$ T_1 $end:math:text$.  
Units lie on a spatial lattice, and outcomes follow a spatial autoregressive structure with spillover from treated units.

Key features of the DGP:

- Spatial correlation governed by $begin:math:text$ \\rho $end:math:text$
- Spillover weights $begin:math:text$ w $end:math:text$ and adjacency matrix $begin:math:text$ W $end:math:text$
- One or more covariates (dimension $begin:math:text$K$end:math:text$)
- A fixed set of treated units
- MCMC estimation with diffuse priors

# Running the Monte Carlo Study

The command below runs multiple scenarios over a grid of $begin:math:text$ (N, T_0, \\rho) $end:math:text$.  
Each scenario involves repeated sampling and estimation.

```{r}
make_rook_W <- function(nrow, ncol, normalize = FALSE) {
  stopifnot(nrow >= 1L, ncol >= 1L)
  idx <- matrix(seq_len(nrow * ncol), nrow, ncol, byrow = TRUE)
  N <- nrow * ncol
  W <- matrix(0, N, N)

  for (r in 1:nrow) {
    for (c in 1:ncol) {
      i <- idx[r, c]
      if (r > 1) {
        W[i, idx[r - 1, c]] <- 1
      }
      if (r < nrow) {
        W[i, idx[r + 1, c]] <- 1
      }
      if (c > 1) {
        W[i, idx[r, c - 1]] <- 1
      }
      if (c < ncol) W[i, idx[r, c + 1]] <- 1
    }
  }

  if (normalize) {
    rs <- rowSums(W)
    W <- W / pmax(rs, 1) # 0割回避
  }
  W
}

run_scenario_many <- function(
  n_sims,
  dgp_args,
  seeds = NULL,
  M = 5000,
  burn = 2000,
  step_rho = 0.02
) {
  res <- run_many_sim(
    n_sims = n_sims,
    dgp_args = dgp_args,
    seeds = seeds,
    M = M,
    burn = burn,
    step_rho = step_rho
  )
  list(
    summary = summarize_many(res),
    raw = res
  )
}

mc_grid_study <- function(
  Ns,
  T0s,
  rhos,
  T1,
  sims_per = 100,
  K = 1,
  beta = c(1.0),
  sigma2 = 1.0,
  treated_idx = 1:4,
  M = 5000,
  burn = 2000,
  step_rho = 0.02,
  seeds = NULL
) {
  # グリッド展開
  grid <- expand.grid(N = Ns, T0 = T0s, rho = rhos, stringsAsFactors = FALSE)

  # 乱数種
  if (is.null(seeds)) {
    seeds <- sample.int(.Machine$integer.max, nrow(grid) * sims_per)
  }
  # シナリオごとの種をスライス
  get_seeds_i <- function(i) {
    idx <- ((i - 1L) * sims_per + 1L):(i * sims_per)
    seeds[idx]
  }

  # 実行
  details <- vector("list", nrow(grid))
  summaries <- vector("list", nrow(grid))

  for (i in seq_len(nrow(grid))) {
    N <- grid$N[i]
    T0 <- grid$T0[i]
    rho <- grid$rho[i]

    # N は正方格子前提（論文の体裁を踏襲）。
    m <- round(sqrt(N))
    if (m * m != N) {
      stop(sprintf(
        "N=%d は完全平方数ではありません（%dx%dにできません）。",
        N,
        m,
        m
      ))
    }

    # W と w
    W <- make_rook_W(m, m)
    w <- numeric(N)
    w[intersect(treated_idx, seq_len(N))] <- 1

    # 真の α
    alpha_true <- rep(0, N)
    alpha_true[1] <- 0.5
    alpha_true[2] <- -0.2
    alpha_true[3:4] <- 0.4
    alpha_true[5:10] <- 0.1 / 6

    # DGP 引数
    dgp_args <- list(
      T0 = T0,
      T1 = T1,
      N = N,
      W = W,
      w = w,
      rho = rho,
      sigma2 = sigma2,
      alpha = alpha_true,
      K = K,
      beta = beta
    )

    # 多回実行
    process_time <- system.time({
      out <- run_scenario_many(
        n_sims = sims_per,
        dgp_args = dgp_args,
        seeds = get_seeds_i(i),
        M = M,
        burn = burn,
        step_rho = step_rho
      )
    })
    print(process_time)

    # シナリオ情報を列として付与
    smry <- out$summary
    smry$N <- N
    smry$T0 <- T0
    smry$T1 <- T1
    smry$rho <- rho
    # 列順を調整（読みやすさ）
    smry <- smry[, c(
      "N",
      "T0",
      "T1",
      "rho",
      "method",
      "bias_point",
      "rmse_point",
      "cover95_point"
    )]

    details[[i]] <- out$raw
    summaries[[i]] <- smry
  }

  list(
    summary = do.call(rbind, summaries),
    details = details,
    grid = grid
  )
}
```


```{r run-grid}
output_dir <- "inst/tables/mc_result"
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

for (N in c(64)) {
  # c(16, 36, 64)
  for (T0 in c(20)) {
    # c(20, 50)
    study <- mc_grid_study(
      Ns = c(N), # grid size: 4×4 and 6×6
      T0s = c(T0), # pre-period length
      rhos = c(-0.8, -0.3, 0.0, 0.3, 0.8), # spatial coefficient c(-0.8, -0.3, 0.0, 0.3, 0.8)
      T1 = 20, # post-treatment length (20 or 50)
      sims_per = 1000, # repetitions per scenario
      K = 1, # number of covariates
      beta = c(1.0), # true regression slope
      sigma2 = 0.1, # innovation variance
      treated_idx = 1:4, # treated units
      M = 5000, # MCMC iterations
      burn = 1000, # burn-in
      step_rho = 0.05 # MH step size
    )
    output_path <- paste0(output_dir, "/mc_study_N=", N, "T0=", T0, ".csv")
    write.csv(study$summary, output_path, row.names = FALSE)
    rm(study)
  }
}

```

The output table contains, for each scenario and estimation method:

- `bias_point`: Bias of the point estimate
- `rmse_point`: Root mean squared error
- `cover95_point`: Coverage of the nominal 95% interval

# Bias and RMSE Plots

```{r plot-bias-rmse, fig.width=7, fig.height=5}
# 論文体裁のパネル表を生成（LaTeX向け）
# 表示名を統一（SCSPILL -> Proposed）し、順序を固定
normalize_methods <- function(df) {
  df %>%
    mutate(
      method = recode(method, SCSPILL = "Proposed"),
      method = factor(method, levels = c("Proposed", "SCM", "BSCM")),
      N = as.integer(N)
    )
}

# 欠け組を安全に補完（NA 埋め）
complete_grid <- function(df) {
  df %>%
    complete(
      N,
      T0,
      T1,
      rho,
      method = factor(
        c("Proposed", "SCM", "BSCM"),
        levels = c("Proposed", "SCM", "BSCM")
      ),
      fill = list(bias_point = NA_real_, rmse_point = NA_real_)
    )
}

# 内部ユーティリティ: wide への整形（列名を "{N}:{method}" に固定）
pivot_metric <- function(df, metric) {
  df %>%
    select(N, T0, T1, rho, method, value = .data[[metric]]) %>%
    pivot_wider(
      names_from = c(N, method),
      values_from = value,
      names_glue = "{N}:{method}"
    )
}

stack_bias_rmse <- function(wide_bias, wide_rmse, rho_order, N_levels) {
  col_keys <- expand.grid(
    N = N_levels,
    method = c("Proposed", "SCM", "BSCM"),
    stringsAsFactors = FALSE
  ) %>%
    mutate(key = paste0(N, ":", method)) %>%
    pull(key)

  pick <- intersect(col_keys, names(wide_bias))
  wide_bias <- wide_bias %>%
    mutate(rho = factor(rho, levels = rho_order)) %>%
    arrange(rho) %>%
    select(rho, all_of(pick))

  wide_rmse <- wide_rmse %>%
    mutate(rho = factor(rho, levels = rho_order)) %>%
    arrange(rho) %>%
    select(rho, all_of(pick))

  bias_block <- wide_bias %>%
    mutate(Measure = "Bias") %>%
    relocate(Measure, .before = rho)
  rmse_block <- wide_rmse %>%
    mutate(Measure = "RMSE") %>%
    relocate(Measure, .before = rho)

  list(
    bias = bias_block,
    rmse = rmse_block,
    col_keys = pick,
    N_levels = N_levels
  )
}

# 1 パネル（特定の T0）を作成
make_panel_table <- function(
  df,
  T0_target,
  rho_order = c(-0.8, -0.3, -0.1, 0.0, 0.1, 0.3, 0.8),
  caption = NULL
) {
  df1 <- df %>% filter(T0 == T0_target)

  N_levels <- sort(unique(df1$N))

  wb <- pivot_metric(df1, "bias_point")
  wr <- pivot_metric(df1, "rmse_point")

  blocks <- stack_bias_rmse(wb, wr, rho_order, N_levels = N_levels)
  bias_block <- blocks$bias
  rmse_block <- blocks$rmse
  key_cols <- blocks$col_keys

  panel_df <- bind_rows(bias_block, rmse_block)

  bottom_labels <- c(
    "Measure",
    "$\\rho$",
    rep(c("Proposed", "SCM", "BSCM"), length(c(16, 36, 64)))
  )
  spanners <- setNames(
    rep(3, length(N_levels)),
    paste0("$N=", N_levels, "$")
  )

  # テーブル本体
  kb <- panel_df %>%
    mutate(across(-c(Measure, rho), ~ ifelse(is.na(.x), NA, round(.x, 3)))) %>%
    kable(
      format = "latex",
      booktabs = TRUE,
      escape = FALSE,
      col.names = c(
        "Measure",
        "$\\rho$",
        rep(c("Proposed", "SCM", "BSCM"), length(N_levels))
      ),
      align = c("l", "c", rep("r", length(key_cols))),
      caption = caption
    ) %>%
    add_header_above(c(" " = 2, spanners)) %>%
    kable_styling(latex_options = c("hold_position")) %>%
    row_spec(
      nrow(bias_block),
      extra_latex_after = "\\midrule[.2pt] \\addlinespace[0.25em]"
    )

  kb
}
```

# Render Panel

```{r plot-coverage, fig.width=7, fig.height=4}
base_dir <- "inst/tables/mc_result"
study <- data.frame()
for (N in c(16, 36, 64)) {
  for (T0 in c(20, 50)) {
    output_path <- paste0(output_dir, "/mc_study_N=", N, "T0=", T0, ".csv")
    df_tmp <- read.csv(output_path)
    study <- study %>% rbind(df_tmp)
  }
}

df_ready <- study %>%
  normalize_methods() %>%
  complete_grid()

spanners <- c("Baseline" = 3, "Follow-up" = 2) # 例：計5列がスパナー対象
tab_a <- make_panel_table(
  df_ready,
  T0_target = 20,
  caption = "Panel (a): $T_0=20$"
)
tab_a
```



```{r}
# Save the LaTeX code for the main panel to a .tex file for inclusion in the manuscript.
table_tex <- capture.output(tab_a, collapse = "\n")
writeLines(table_tex, "./inst/tables/simulation_tables.tex")
```

```{r}
sessionInfo()
```