---
title: "Monte Carlo Simulation Study for Spatial Spillover Synthetic Control"
author: "Replication Package"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

# Overview

This vignette reproduces the finite‐sample evaluation of the spatial spillover synthetic control estimator.  
We conduct a Monte Carlo study under controlled data‐generating processes (DGPs) with spatial dependence and interference.

The following functions are assumed to be already installed and available (part of the replication package):

- `run_many_sim()`: Run repeated simulations under one scenario  
- `summarize_many()`: Summarize bias, RMSE, and 95% coverage  
- `make_rook_W()`: Generate rook‐adjacency matrix on an m×n grid  
- `run_scenario_many()`: Convenience wrapper for one scenario  
- `mc_grid_study()`: Grid simulation across (N, T0, ρ) values  

```{r setup, message=FALSE, warning=FALSE}
library(scspill)
library(dplyr)
library(tidyr)
library(ggplot2)

# Assume the simulation engine functions are already part of the package
set.seed(20251030)
```

# Data Generating Process

We simulate an untreated pre‐period of length $begin:math:text$ T_0 $end:math:text$ and a post‐period of length $begin:math:text$ T_1 $end:math:text$.  
Units lie on a spatial lattice, and outcomes follow a spatial autoregressive structure with spillover from treated units.

Key features of the DGP:

- Spatial correlation governed by $begin:math:text$ \\rho $end:math:text$
- Spillover weights $begin:math:text$ w $end:math:text$ and adjacency matrix $begin:math:text$ W $end:math:text$
- One or more covariates (dimension $begin:math:text$K$end:math:text$)
- A fixed set of treated units
- MCMC estimation with diffuse priors

# Running the Monte Carlo Study

The command below runs multiple scenarios over a grid of $begin:math:text$ (N, T_0, \\rho) $end:math:text$.  
Each scenario involves repeated sampling and estimation.

```{r}
make_rook_W <- function(nrow, ncol, normalize = FALSE) {
  stopifnot(nrow >= 1L, ncol >= 1L)
  idx <- matrix(seq_len(nrow * ncol), nrow, ncol, byrow = TRUE)
  N <- nrow * ncol
  W <- matrix(0, N, N)

  for (r in 1:nrow) {
    for (c in 1:ncol) {
      i <- idx[r, c]
      if (r > 1) {
        W[i, idx[r - 1, c]] <- 1
      }
      if (r < nrow) {
        W[i, idx[r + 1, c]] <- 1
      }
      if (c > 1) {
        W[i, idx[r, c - 1]] <- 1
      }
      if (c < ncol) W[i, idx[r, c + 1]] <- 1
    }
  }

  if (normalize) {
    rs <- rowSums(W)
    W <- W / pmax(rs, 1) # 0割回避
  }
  W
}

run_scenario_many <- function(
  n_sims,
  dgp_args,
  seeds = NULL,
  M = 5000,
  burn = 2000,
  step_rho = 0.02
) {
  res <- run_many_sim(
    n_sims = n_sims,
    dgp_args = dgp_args,
    seeds = seeds,
    M = M,
    burn = burn,
    step_rho = step_rho
  )
  list(
    summary = summarize_many(res),
    raw = res
  )
}

mc_grid_study <- function(
  Ns,
  T0s,
  rhos,
  T1,
  sims_per = 100,
  K = 1,
  beta = c(1.0),
  sigma2 = 1.0,
  treated_idx = 1:4,
  M = 5000,
  burn = 2000,
  step_rho = 0.02,
  seeds = NULL
) {
  # グリッド展開
  grid <- expand.grid(N = Ns, T0 = T0s, rho = rhos, stringsAsFactors = FALSE)

  # 乱数種
  if (is.null(seeds)) {
    seeds <- sample.int(.Machine$integer.max, nrow(grid) * sims_per)
  }
  # シナリオごとの種をスライス
  get_seeds_i <- function(i) {
    idx <- ((i - 1L) * sims_per + 1L):(i * sims_per)
    seeds[idx]
  }

  # 実行
  details <- vector("list", nrow(grid))
  summaries <- vector("list", nrow(grid))

  for (i in seq_len(nrow(grid))) {
    N <- grid$N[i]
    T0 <- grid$T0[i]
    rho <- grid$rho[i]

    # N は正方格子前提（論文の体裁を踏襲）。
    m <- round(sqrt(N))
    if (m * m != N) {
      stop(sprintf(
        "N=%d は完全平方数ではありません（%dx%dにできません）。",
        N,
        m,
        m
      ))
    }

    # W と w
    W <- make_rook_W(m, m)
    w <- numeric(N)
    w[intersect(treated_idx, seq_len(N))] <- 1

    # 真の α
    alpha_true <- rep(0, N)
    alpha_true[1] <- 0.5
    alpha_true[2] <- -0.2
    alpha_true[3:4] <- 0.4
    alpha_true[5:10] <- 0.1 / 6

    # DGP 引数
    dgp_args <- list(
      T0 = T0,
      T1 = T1,
      N = N,
      W = W,
      w = w,
      rho = rho,
      sigma2 = sigma2,
      alpha = alpha_true,
      K = K,
      beta = beta
    )

    # 多回実行
    process_time <- system.time({
      out <- run_scenario_many(
        n_sims = sims_per,
        dgp_args = dgp_args,
        seeds = get_seeds_i(i),
        M = M,
        burn = burn,
        step_rho = step_rho
      )
    })
    print(process_time)

    # シナリオ情報を列として付与
    smry <- out$summary
    smry$N <- N
    smry$T0 <- T0
    smry$T1 <- T1
    smry$rho <- rho
    # 列順を調整（読みやすさ）
    smry <- smry[, c(
      "N",
      "T0",
      "T1",
      "rho",
      "method",
      "bias_point",
      "rmse_point",
      "cover95_point"
    )]

    details[[i]] <- out$raw
    summaries[[i]] <- smry
  }

  list(
    summary = do.call(rbind, summaries),
    details = details,
    grid = grid
  )
}
```


```{r run-grid}
study <- mc_grid_study(
  Ns = c(16, 36, 64), # grid size: 4×4 and 6×6
  T0s = c(30, 60), # pre-period length
  rhos = c(0.3), # spatial coefficient
  T1 = 20, # post-treatment length (20 or 50)
  sims_per = 10, # repetitions per scenario
  K = 1, # number of covariates
  beta = c(1.0), # true regression slope
  sigma2 = 1.0, # innovation variance
  treated_idx = 1:4, # treated units
  M = 5000, # MCMC iterations
  burn = 1000, # burn-in
  step_rho = 0.15 # MH step size
)

head(study$summary, 10)
```

The output table contains, for each scenario and estimation method:

- `bias_point`: Bias of the point estimate
- `rmse_point`: Root mean squared error
- `cover95_point`: Coverage of the nominal 95% interval

# Bias and RMSE Plots

```{r plot-bias-rmse, fig.width=7, fig.height=5}
# ---------- Helpers: build table blocks (Bias / RMSE) ----------
build_block_wide <- function(
  df_panel,
  metric = c("bias", "rmse"),
  methods_order = c("Proposed", "SCM", "BSCM"),
  digits = 3
) {
  metric <- match.arg(metric)
  value_col <- if (metric == "bias") "bias_point" else "rmse_point"

  agg <- df_panel %>%
    group_by(rho, N, method) %>%
    summarise(value = mean(.data[[value_col]], na.rm = TRUE), .groups = "drop")

  Ns <- sort(unique(agg$N))
  agg$method <- factor(agg$method, levels = methods_order)

  wide <- agg %>%
    mutate(col = paste0("N", N, "::", as.character(method))) %>%
    select(rho, col, value) %>%
    tidyr::pivot_wider(names_from = col, values_from = value) %>%
    arrange(rho)

  num_cols <- setdiff(colnames(wide), "rho")
  wide[num_cols] <- lapply(wide[num_cols], function(x) {
    sprintf(paste0("%.", digits, "f"), x)
  })

  col_order <- unlist(lapply(Ns, function(n) {
    paste0("N", n, "::", methods_order)
  }))
  wide <- wide[, c("rho", col_order), drop = FALSE]
  attr(wide, "Ns") <- Ns
  attr(wide, "methods_order") <- methods_order
  wide
}

# ---------- Combine Bias + RMSE with a dashed separator ----------
combine_bias_rmse <- function(bias_wide, rmse_wide) {
  stopifnot(identical(colnames(bias_wide), colnames(rmse_wide)))
  common_cols <- colnames(bias_wide)

  out <- dplyr::bind_rows(
    bias_wide,
    setNames(as.list(rep("—", length(common_cols))), common_cols),
    rmse_wide
  )
  out
}

# ---------- Two-row headers for kable ----------
make_kable_header <- function(wide_df) {
  Ns <- attr(wide_df, "Ns")
  methods <- attr(wide_df, "methods_order")

  top <- c(" " = 1) # for 'Block'/'rho' columns later
  for (n in Ns) {
    top[paste0("$N=", n, "$")] <- length(methods)
  }

  second <- c(" $\\rho$ " = 1)
  for (n in Ns) {
    second <- c(second, setNames(rep(methods, 1), methods))
  }
  list(top = top, second = second)
}
```

# Render Panel

```{r plot-coverage, fig.width=7, fig.height=4}
render_panel_table <- function(
  summary_df,
  T0_target,
  methods_order = c("Proposed", "SCM", "BSCM"),
  digits = 3,
  caption = NULL
) {
  df_panel <- summary_df %>% filter(T0 == T0_target)
  stopifnot(nrow(df_panel) > 0)

  bias_w <- build_block_wide(
    df_panel,
    metric = "bias",
    methods_order = methods_order,
    digits = digits
  )
  rmse_w <- build_block_wide(
    df_panel,
    metric = "rmse",
    methods_order = methods_order,
    digits = digits
  )
  merged <- combine_bias_rmse(bias_w, rmse_w)

  hdr <- make_kable_header(bias_w)

  # locate dashed-separator row (filled with em-dashes)
  dash_row <- which(merged[[1]] == "—")
  if (length(dash_row) == 1L) {
    merged[dash_row, ] <- lapply(merged[dash_row, ], function(x) "")
  }

  # prepend block labels: 'Bias' and 'RMSE' stubs
  rhos <- sort(unique(df_panel$rho))
  n_rho <- length(rhos)
  block_label <- c(rep("Bias", n_rho), "", rep("RMSE", n_rho))
  merged <- cbind(Block = block_label, merged)

  kb <- kable(
    merged,
    format = "latex",
    booktabs = TRUE,
    escape = FALSE,
    caption = caption,
    align = c("l", "r", rep("r", ncol(merged) - 2))
  ) %>%
    kableExtra::kable_styling(latex_options = c("hold_position")) %>%
    add_header_above(hdr$second, escape = FALSE) %>%
    add_header_above(hdr$top, escape = FALSE)

  if (length(dash_row) == 1L) {
    kb <- kb %>%
      kableExtra::row_spec(dash_row + 1, extra_latex_after = "\\hdashline")
  }
  kb
}
```

# Inspecting Raw Output for One Scenario

```{r inspect-scenario}
unique_T0 <- sort(unique(study$summary$T0))
panel_caption <- paste0(
  "Simulation Results for Bias and RMSE \\quad (Panel (a): $T_0 = ",
  unique_T0[1],
  "$)"
)

panel_a <- render_panel_table(
  summary_df = study$summary,
  T0_target = unique_T0[1],
  methods_order = c("Proposed", "SCM", "BSCM"),
  digits = 3,
  caption = panel_caption
)

panel_a
```

```{r}
if (length(unique_T0) >= 2) {
  for (j in 2:length(unique_T0)) {
    cat(sprintf("**Panel (%s):** $T_0 = %d$\n\n", letters[j], unique_T0[j]))
    print(
      render_panel_table(
        summary_df = study$summary,
        T0_target = unique_T0[j],
        methods_order = c("Proposed", "SCM", "BSCM"),
        digits = 3,
        caption = paste0(
          "Simulation Results for Bias and RMSE \\quad (Panel (",
          letters[j],
          "): $T_0 = ",
          unique_T0[j],
          "$)"
        )
      )
    )
    cat("\n\n")
  }
}
```

```{r}
# Save the LaTeX code for the main panel to a .tex file for inclusion in the manuscript.
save_kable(panel_a, file = "./data/table_sim_panel_a.tex")
```

```{r}
sessionInfo()
```