---
title: "Geweke (2004) Joint Distribution Test for the SC-Spillover Sampler"
output: html_document
---

This vignette implements the **Joint Distribution Test (JDT)** of Geweke (2004) to formally validate the posterior transition kernel used in the SC-spillover MCMC sampler. The JDT checks whether the **successive-conditional (SC)** simulator — i.e., one-step posterior kernel updates — generates draws consistent with the same **joint distribution** \(p(\theta, y) = p(y \mid \theta)\pi(\theta)\) as the **marginal-conditional (MC)** simulator, which draws \(\theta \sim \pi(\theta)\) independently and then simulates \(y \sim p(y \mid \theta)\).

If both simulators are correct and share the same prior and likelihood,  
the expectation of any integrable statistic \(g(\theta, y)\) must coincide under both paths.  
The JDT forms a Z-statistic comparing the two Monte Carlo estimates.

---

## Setup

We create a simulated spatial–spillover synthetic control environment with covariates and one latent factor. The forward data generating mechanism matches the structural SAR spillover model used by the MCMC kernel.

```{r}
rm(list = ls())
# library(scspill)
library(dplyr)
library(tidyr)
library(stringr)
library(knitr)
library(kableExtra)

source("R/01_utils.R")
source("R/10_sc_spillover.R")
source("R/21_mcmc_alpha.R")
source("R/22_mcmc_sar.R")
source("R/30_simulation_.R")
source("R/40_geweke_latest.R")
source("R/41_robustness_check.R")
Rcpp::sourceCpp("src/20_mcmc.cpp")
Rcpp::sourceCpp("src/40_geweke_latest.cpp")
```

```{r}
# --- dimensions ---
T0 <- 15
N <- 8
K <- 2 # X covariates
p <- 1 # latent factor

# --- W, w, alpha_hat_scaled ---
W_raw <- matrix(0, N, N)
for (i in 1:N) {
  if (i > 1) {
    W_raw[i, i - 1] <- 1
  }
  if (i < N) W_raw[i, i + 1] <- 1
}
w_raw <- rep(0, N)
w_raw[1] <- 1

W_use <- row_normalize(W_raw)
w_use <- as.numeric(w_raw)
wsum <- sum(w_use)
if (is.finite(wsum) && wsum > 1e-12) {
  w_use <- w_use / wsum
}

# fixed alpha_hat_scaled (in practice, scale BSCM alpha)
alpha_hat_scaled <- rnorm(N, 0, 0.4)

# --- X: T0 x N x K ---
Xc_pre <- array(0, dim = c(T0, N, K))
for (t in 1:T0) {
  for (i in 1:N) {
    for (k in 1:K) {
      Xc_pre[t, i, k] <- rnorm(1)
    }
  }
}

# --- true parameters (for internal consistency) ---
rho_true <- 0.25
sigma2_true <- 0.5^2
beta_true <- c(0.7, -0.5)
Eta_true <- matrix(rnorm(N * p, 0, 0.5), N, p)
Gamma_true <- matrix(0, p, T0)
phi_true <- 0.5
s2g_true <- 0.2
for (t in 1:T0) {
  if (t == 1) {
    Gamma_true[, t] <- rnorm(p, 0, sqrt(s2g_true / (1 - phi_true^2)))
  }
  if (t >= 2) {
    Gamma_true[, t] <- phi_true *
      Gamma_true[, t - 1] +
      rnorm(p, 0, sqrt(s2g_true))
  }
}

# --- prior ---
a0 <- 1.0
b0 <- 1.0 # IG(a0, b0) prior for sigma2
step_rho <- 0.05

# --- forward simulator in R (mirror of C++ forward kernel) ---
sim_forward_R <- function(
  T0,
  W_use,
  w_use,
  alpha_hat_scaled,
  rho,
  sigma2,
  Xc_pre,
  beta,
  Eta,
  Gamma
) {
  N <- nrow(W_use)
  K <- dim(Xc_pre)[3]
  A <- W_use + w_use %*% t(alpha_hat_scaled)
  I <- diag(N)
  Yc <- matrix(NA_real_, T0, N)
  for (t in 1:T0) {
    mu <- rep(0, N)
    if (K > 0) {
      Xt <- Xc_pre[t, , ] # N x K
      mu <- mu + as.vector(Xt %*% beta)
    }
    if (ncol(Eta) > 0) {
      mu <- mu + as.vector(Eta %*% Gamma[, t])
    }
    eps <- rnorm(N, 0, sqrt(sigma2))
    rhs <- mu + eps
    M <- I - rho * A
    Yc[t, ] <- as.vector(solve(M, rhs))
  }
  Yc
}

# pseudo observed treated pre-outcome (for g-function)
Y0_pre <- rnorm(T0)
```

---

## Running the Joint Distribution Test

The function `geweke_jdt_full()` executes the marginal–conditional (MC) and successive-conditional (SC) simulators and reports Z-statistics comparing the expectations of user-defined statistics \(g(\theta,y)\).

```{r}
# JDT parameters
M1 <- 2000000L # MC path draws
M2 <- 2000000L # SC path draws
burn_in <- 2000000L

out <- geweke_jdt_full(
  Y0_pre = Y0_pre,
  Yc_pre_like_dims = c(T0, N),
  W = W_use,
  w = w_use,
  alpha_hat_scaled = alpha_hat_scaled,
  Xc_pre = Xc_pre,
  p = p,
  M1 = M1,
  M2 = M2,
  burn_in = burn_in,
  a0 = a0,
  b0 = b0,
  step_rho = step_rho,
  g_fn = default_g_fn,
  verbose = TRUE
)

print(out$summary)

write.csv(out$summary, "inst/tables/geweke_jdt_summary.csv", row.names = FALSE)
```

```{r}
jdt_tbl <- out$summary

row_renamer <- c(
  rho = "$\\rho$",
  log_sigma2 = "$\\log(\\sigma^2)$",
  yc_mean = "$\\bar{y}_c$",
  log_yc_var = "$\\log(\\mathrm{Var}(y_c))$",
  spatial_quadratic = "$h_{\\text{spatial}}$",
  corr_y0_wyc = "$\\mathrm{Corr}(y_0, Wy)$",
  beta_mean = "$\\bar{\\beta}$",
  Eta_mean = "$\\bar{\\eta}$",
  Gamma_mean = "$\\bar{\\Gamma}$"
)

jdt_tbl$g <- unname(ifelse(
  jdt_tbl$g %in% names(row_renamer),
  row_renamer[jdt_tbl$g],
  jdt_tbl$g
))

colnames(jdt_tbl) <- c(
  "Statistic $g$",
  "Mean (iid)",
  "Mean (MCMC)",
  "SE (iid)",
  "SE (MCMC)",
  "$Z$",
  "$p$-value"
)

knitr::kable(jdt_tbl, caption = "Summary of Joint Distribution Test Results")

jdt_tex <- kableExtra::kable(
  jdt_tbl,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "Summary of Joint Distribution Test Results"
)
cat(jdt_tex, file = "./inst/tables/geweke_jdt_summary.tex")
```

```{r}
thin_app <- 5L
out_app_thin <- geweke_jdt_full(
  Y0_pre = Y0_pre,
  Yc_pre_like_dims = c(T0, N),
  W = W_use,
  w = w_use,
  alpha_hat_scaled = alpha_hat_scaled,
  Xc_pre = Xc_pre,
  p = p,
  M1 = M1, # MC path unchanged
  M2 = M2 * thin_app, # run longer then thin inside g_fn if you implement that
  burn_in = burn_in,
  a0 = a0,
  b0 = b0,
  step_rho = 0.05,
  g_fn = default_g_fn,
  verbose = FALSE,
  rho_support = NULL
)

# (b) Custom g-function (example): centered squared norm of Y and rho^2
g_fn_custom <- function(theta, Y, W, w, Y0_pre) {
  yc <- as.numeric(Y)
  c(
    rho = theta$rho,
    rho_sq = theta$rho^2,
    log_sigma2 = log(theta$sigma2 + 1e-12),
    yc_mean = mean(yc),
    yc_center_q = mean((yc - mean(yc))^2)
  )
}

out_app_custom <- geweke_jdt_full(
  Y0_pre = Y0_pre,
  Yc_pre_like_dims = c(T0, N),
  W = W_use,
  w = w_use,
  alpha_hat_scaled = alpha_hat_scaled,
  Xc_pre = Xc_pre,
  p = p,
  M1 = M1,
  M2 = M2,
  burn_in = burn_in,
  a0 = a0,
  b0 = b0,
  step_rho = 0.05,
  g_fn = g_fn_custom, # <- custom statistics
  verbose = FALSE,
  rho_support = NULL
)

knitr::kable(
  out_app_custom$summary,
  caption = "Appendix JDT with custom g-function."
)
```

---

## Interpretation

- Each row corresponds to one statistic \(g(\theta, y)\): e.g., \( \rho\), \(\log\sigma^2\), mean outcomes, spatial quadratic forms, etc.
- Under correct implementation, **Z-statistics should be close to zero**, and p-values should not systematically reject.
- Strong deviations suggest **mismatch between the MC and SC simulators**, e.g.:
  - prior inconsistency,
  - incorrect acceptance/rejection logic,
  - wrong support for \(\rho\),
  - or incorrect likelihood in the forward simulation step.

> **Important:**  
> JDT requires **exact equality** of the prior, likelihood, and support used in the MC simulator and those implicitly assumed in the MCMC transition kernel. Any discrepancy will — correctly — trigger rejections.

---

## Practical Guidance

- Increase `M1`, `M2`, and `burn_in` for robustness, especially with spatial dependence and latent factors.
- Tune `step_rho` to avoid extremely low or high Metropolis acceptance.
- Ensure `Xc_pre` remains a 3-D array `(T0 × N × K)`; set `drop = FALSE` when slicing.
- `alpha_hat_scaled` must match the normalization used in the BSCM preprocessing.

---

## References

- Geweke, J. (2004), *Getting it Right: Joint Distribution Tests of Posterior Simulators*.  
  Journal of the American Statistical Association.